---
title: "46-926 Homework 2, Part I"
author: "Jingyi Guo, Pittsburgh Campus"
date: "1/28/2017"
output: pdf_document
---

```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE,warning=FALSE}
library(leaps)
library(bestglm)
library(MASS)
library(car)
bonddata = read.table("http://www.stat.cmu.edu/~cschafer/MSCF/bonddata.txt",sep=",",header=T)
newdata=subset(bonddata[3:16])
trade_price=newdata$trade_price
weight=newdata$weight
current_coupon=newdata$current_coupon
time_to_maturiy=newdata$time_to_maturiy
is_callable=newdata$is_callable
reporting_delay=newdata$reporting_delay
trade_size=newdata$trade_size
trade_type=newdata$trade_type
curved_based_price=newdata$curve_based_price
received_time_diff_last1=newdata$received_time_diff_last1
trade_price_last1=newdata$trade_price_last1
trade_size_last1=newdata$trade_size_last1
trade_type_last1=newdata$trade_type_last1
curved_based_price_last1=newdata$curve_based_price_last1
```

## 1 Exhaustive Search

```{r 1}
allpreds=cbind(weight,current_coupon,time_to_maturiy,is_callable,reporting_delay,trade_size,trade_type,
               curved_based_price,received_time_diff_last1,trade_price_last1,trade_size_last1,
               trade_type_last1,curved_based_price_last1)
Xyframe=data.frame(cbind(allpreds,trade_price))
bestmod=bestglm(Xyframe,IC="AIC")
print(bestmod)
```

We see that the final model takes the form

\[Y_i=\beta_0+\beta_1trade\_size+\beta_2trade\_type+\beta_3curved\_based\_price+\beta_4received\_time\_diff\_last1+\beta_5trade\_price\_last1\]
\[\ +\beta_6trade\_type\_type\_last1+\beta_7curved\_based\_price\_last1\]

Categorical predictors trade_type, trade_type_last1 appear in the final model.

## 2 PRESS

First, compute the PRESS for the full model

```{r 2 full PRESS}
fitfullmodel = lm(trade_price ~ ., data = newdata)
levs=hatvalues(fitfullmodel)
PRESSfull=sum((fitfullmodel$resid/(1-levs))^2)
print(PRESSfull)
```

Then, compute the PRESS for the final model in Question 1

```{r 2 Q1 PRESS}
fitmodel1 = lm(trade_price ~ trade_type+curved_based_price+curved_based_price+received_time_diff_last1
               +trade_price_last1+trade_type_last1+curved_based_price_last1, data = newdata)
levs1=hatvalues(fitmodel1)
PRESS1=sum((fitmodel1$resid/(1-levs1))^2)
print(PRESS1)
```

The PRESS value for the full model is larger than that for the AIC-optimal value found in Question 1. So according to PRESS, the model in Question 1 has higher predictive power than the full model.


## 3 Influential Observations

```{r 3}
cookd=as.numeric(cooks.distance(fitmodel1))
sort(pf(cookd,8,1612),decreasing=TRUE)[1:5]
```

The largest Cook's Distance is at the 3.0% of the F distribution. So there is no reason for concern from influcential observations.

## 4 Robust Regression Method

```{r 4}
holdrlm1 = rlm(trade_price ~ trade_type+curved_based_price+curved_based_price+received_time_diff_last1
               +trade_price_last1+trade_type_last1+curved_based_price_last1, data = newdata)
summary(holdrlm1)
```

## 5 Box-Cox Procedure

```{r 5}
boxcox(trade_price ~ .,data = newdata)
```

From the plot, 1 is very close to being in the 95% Confidence Interval of lambda. So we could approximate lambda to be 1, which means a transformation of response is not necessary.

However, to be more precise, we see that the optimal choice is \[\lambda\approx1.1.\] And this can be applied to the data as follows
```{r 5 transformation}
transrealizedvol=bcPower(trade_price,1.1)
```

