---
title: "46-926 Homework 5, Part 1"
author: "Jingyi Guo, Pittsburgh Campus"
date: "2/18/2017"
output: pdf_document
---

## Set Up

```{r set up,warning=FALSE,message=FALSE,results='hide'}
#set up
library(mgcv)
bonddata = read.table("http://www.stat.cmu.edu/~cschafer/MSCF/bonddata.txt",sep=",",header=T)
bonddatasub = bonddata[,-c(1,2,17:61)]
#convert the factors
bonddatasub$is_callable = factor(bonddatasub$is_callable)
bonddatasub$trade_type = factor(bonddatasub$trade_type)
bonddatasub$trade_type_last1 = factor(bonddatasub$trade_type_last1)
```

## 1 Transformation

```{r 1}
bonddatasub$weight=log(bonddatasub$weight)
bonddatasub$time_to_maturity=log(bonddatasub$time_to_maturity)
bonddatasub$trade_size=log(bonddatasub$trade_size)
bonddatasub$trade_size_last1=log(bonddatasub$trade_size_last1)
#transform to categorical variable
bonddatasub$reporting_delay=cut(bonddatasub$reporting_delay,c(-Inf,2,10,100,Inf))
bonddatasub$received_time_diff_last1=cut(bonddatasub$received_time_diff_last1,c(-Inf,500,
                                                                                75000,4000000,Inf))

```

## 2 GAM Model

```{r 2,warning=FALSE}
holdgam=gam(trade_price ~ s(weight)+s(current_coupon)+s(time_to_maturity)+s(trade_size)
            +s(curve_based_price)+s(trade_price_last1)+s(trade_size_last1)+s(curve_based_price_last1)
+is_callable+reporting_delay+trade_type+received_time_diff_last1+trade_type_last1,
data=bonddatasub)
summary(holdgam)
```

## 3 

From the summary, we can see that the model predicts the mean difference in trade price between bonds whose current trade is of type “3” and a bond whose current trade is of type “4" to be 1.4828-0.7326=0.7502.

Therefore, trade price of type 3 is 0.7502 higher than type 4 on average.

## 4

```{r 4,warning=FALSE}
plot(holdgam, pages=2,scale=0,scheme=1)
```

## 5

```{r 5,warning=FALSE}
plot(holdgam$fit, holdgam$residuals,
     xlab="Fitted Values", ylab="Residuals",pch=16)
```

Quality of the fit: There is no prevalent pattern in the plot of residuals versus fitted values. However, there is one point for which the residual is quite extreme relative to others.

## 6

```{r 6,warning=FALSE}
plot(predict(holdgam),bonddatasub$trade_price,pch=16,cex=0.7,xlab="Predicted Value",
     ylab="Actual Response", cex.axis=1.3,cex.lab=1.3)
abline(0,1,lwd=2,col=2)
```

Quality of the fit: almost all the points are close to the 45 degree line and they are evenly distributed on both sides of the line. However there is a point far from the line worth noticing.

## 7

```{r 7,warning=FALSE}
# The alternative model
holdlinear = gam(trade_price ~ weight + current_coupon +
                   time_to_maturity + is_callable + reporting_delay +
                   trade_size + trade_type + curve_based_price +
                   received_time_diff_last1 + trade_price_last1 +
                   trade_size_last1 + trade_type_last1 + curve_based_price_last1,
                 data = bonddatasub)
#AIC of linear model
AIC(holdlinear)

# AIC of gam model:
AIC(holdgam)
```

AIC of the gam model is smaller than that of linear model. So the extra complexity is justified since it makes AIC smaller.
