---
title: "46-926 Homework 1, Part 1"
author: "Jingyi Guo"
date: "1/22/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Categorical predictors: 
id, bond_id, is_callable, trade_type, trade_type_last1, trade_type_last2, ..., trade_type_last10

exclude categorical predictors and predictors from time steps 5 and prior:
```{r exclude predictors}
bonddata = read.table("http://www.stat.cmu.edu/~cschafer/MSCF/bonddata.txt",sep=",",header=T)
newdata=subset(bonddata[1:31],select=-c(id,bond_id,is_callable,trade_type,trade_type_last1,trade_type_last2,trade_type_last3,trade_type_last4))
```


2.  Fit a model using all of the predictors that remain: 

```{r fit model}
fitmodel = lm(trade_price ~ ., data = newdata)
```

AIC for this model:
```{r AIC}
AIC(fitmodel)
```


Plot of residuals versus fitted values:
```{r plot 1}
plot(as.numeric(fitmodel$fit),as.numeric(fitmodel$resid),pch=16,xlab="Fitted Values", ylab="Residuals")
#changed the scales of axes
plot(as.numeric(fitmodel$fit),as.numeric(fitmodel$resid),pch=16,xlab="Fitted Values", ylab="Residuals",xlim=c(80,140),ylim=c(-10,10))
```

Comment: There is no significantly prevalent pattern in the plot of residual versus fitted values. However, there is one point fot which the residual is quite extreme relative to others.

The normal probability plot of the residuals:
```{r plot 2}
qqnorm(as.numeric(fitmodel$resid),pch=16,main="")
qqline(as.numeric(fitmodel$resid))
```

Comment: The normal probability plot of residuals suggests that errors have distribution with tails significantly heavier than normal distribution.
